A meta analysis of tournaments and an evaluation of performance in the Iterated Prisoner's Dilemma.

### 2019-11-11 paper/main.tex
new draft with additions

\newcommand{\uniquenumberofseeds}{\input{unique_number_of_seeds.tex}}
\newcommand{\numberofalltournaments}{\input{number_of_all_tournaments.tex}}

@@

\begin{abstract}

The Iterated Prisoner's Dilemma has been used for decades as a model of
behavioural interactions. From the celebrated performance of Tit for Tat, to the
introduction of the zero-determinant strategies, to the use of sophisticated
structures such as neural networks, the literature has been exploring the
performance of strategies in the game for years. Most of these strategies are
now accessible due to an open source package; Axelrod-Python. This manuscript
make use of Axelrod-Python to conduct a meta analysis of Iterated Prisoner's
Dilemma tournaments. The aim is to evaluate the performance of 194 strategies
over \numberofalltournaments Iterated Prisoner's Dilemma tournaments and to
explore the factors of their success.
\end{abstract}

@@

Another set of IPD strategies that have received a lot of attention
are the zero-determinants~\cite{Press2012}. By
forcing a linear relationship between the payoffs the zero-determinant
strategies can ensure that they will never receive less than their opponents.
The
American Mathematical Society's news section stated that ``the world of game
theory is currently on fire''.
Zero-determinant strategies are indeed a set of mathematically unique strategies
and robust in pairwise interactions, however, their simplicity and extortionate
behaviour have been tested. In~\cite{Harper2017} a tournament containing over
200 strategies, including zero-determinants, was ran. None of the zero-determinant
strategies ranked in top spots. Instead, the top ranked strategies were a set of
trained strategies based lookup tables~\cite{Axelrod1987}, hidden markov
models~\cite{Harper2017} and finite state automata~\cite{Miller1996}.

Though only a select pieces of work have been discussed, the IPD is
rich, and new strategies and competitions are being published every year.
The question, however, still remains the same: what is the best way to play the
game?

This manuscript uses an open source package called
Axelrod-Python~\cite{axelrodproject} to simulate a large number of computer
tournaments, using as many strategies as possible from the literature. The aim
is to evaluate the performance of 194 strategies in IPD tournaments, and
futhermore, explore the factors of their success. This is done not only for
standard round robin tournaments, but also for tournaments with noise and
tournaments with a probabilistic ending. The different tournament types as well
as the data collection are covered in Section~\ref{section:data_collection}.
Section~\ref{section:top_performances}, focuses on the best performing strategies
for each type of tournament and overall.
Section~\ref{section:evaluation_of_performance}, explores the traits which
contribute to good performance, and finally the results are summarised in
Section~\ref{section:conclusion}.

This manuscripts uses several parameters. These are introduced in the following
sections, however, the full set of parameters and their definitions are given in
Appendix~\ref{app:parameters}. All the data is available at~\cite{data}.

@@

For the purposes of this manuscript a data set containing results of IPD tournaments
has been generated and is available at~\cite{data}. This
was done using the open source package Axelrod-Python~\cite{axelrodproject}, and
more specifically, version 3.0.0. Axelrod-Python allows for different types of
IPD computer tournaments to be simulated whilst
containing a list of over 180 strategies. Most of these are strategies described
in the literature with a few exceptions being strategies that have been
contributed specifically to the package. This paper make use of 194 strategies
implemented in version 3.0.0. A list of the strategies is given in the
Appendix~\ref{app:list_of_players}. Though Axelrod-Python features several
tournament types, this work considers only standard, noisy, probabilistic ending
and noisy probabilistic ending tournaments.

\textbf{Standard tournaments}, are tournaments similar to that of Axelrod's
in~\cite{Axelrod1980a}. There are \(N\) strategies which all play an iterated
game of \(n\) number of turns against each other. Note that self interactions
are not included. Similarly, \textbf{noisy
tournaments} have \(N\) strategies and \(p_n\) number of turns, but at each turn
there is a probability \(p_n\) that a player's action will be flipped.
\textbf{Probabilistic ending tournaments}, are of size \(N\) and after each turn
a match between strategies ends with a given probability \(p_e\). Finally,
\textbf{noisy probabilistic ending} tournaments have both a noise probability
\(p_n\) and an ending probability \(p_e\). For smoothing the simulated results a
tournament is repeated for \(k\) number of times. To evaluate the effect of smoothing
this is allowed to vary. The winner of each tournament
is based on the average score a strategy achieved and not by number of wins.

% A summary of each tournaments' parameters is given in
% Table~\ref{table:tournaments_parameters}.

% \begin{table}[!htbp]
%     \begin{center}
%         \resizebox{.85\textwidth}{!}{
%         \begin{tabular}{lcccccc}
%     \toprule
%     tournament type & number of strategies & repetitions & turns & noise probability & probability of match ending \\
%     \midrule
%     standard & $N$ & k & $n$ & - & - \\
%     noisy & $N$  & k & $n$ & $p$ & - \\
%     probabilistic ending & $N$ & k & - & - & $e$ \\
%     noisy probabilistic ending & $N$ & k & - & $p$ & $e$ \\
%     \bottomrule
%         \end{tabular}}
%     \end{center}
%     \caption{Tournament types' parameters.}
%     \label{table:tournaments_parameters}
% \end{table}

The process of collecting tournament results implemented in this manuscript is described by
Algorithm~\ref{algorithm:data_generation}. For each trial a random size \(N\) is
selected, and from the 194 strategies a random list of \(N\) strategies is
chosen. For the given list of strategies a standard, a noisy, a probabilistic
ending and a noisy probabilistic ending tournament are performed and repeated
\(k\) times. The parameters for the tournaments as well as the number of
repetitions are selected once for each trial. The parameters and their
respective minimum and maximum values are given by
Table~\ref{table:parameters_values}.

@@

The source code for the data collection as well as the source code for
the analysis, which will be discussed in the following sections, have been written
following best practices~\cite{Aberdour2007, Benureau2018}. It has been packaged
and is available here. %TODO archive code

\begin{algorithm}[!htbp]
    \setstretch{1.35}
    \ForEach{\text{seed} $\in [0, 11420]$}{
        $N \gets \text{randomly select integer}\in [N_{min}, N_{max}]$\;
        $\text{players} \gets  \text{randomly select $N$ players}$\;
        $k \gets  \text{randomly select integer}\in [k_{min}, k_{max}]$\;
        $n \gets  \text{randomly select integer}\in [n_{min}, n_{max}]$\;
        $p \gets  \text{randomly select float}\in [p_{n\, min}, p_{n\, max}]$\;
        $e \gets   \text{randomly select float}\in [p_{e\, min}, p_{e\, max}]$\;
        \vspace{0.4cm}
        $\text{result standard}$ $\gets$ Axelrod.tournament$(\text{players}, n, k)$\;
        $\text{result noisy}$ $\gets$ Axelrod.tournament$(\text{players}, n, p_n, k)$\;
        $\text{result probabilistic ending}$ $\gets$ Axelrod.tournament$(\text{players}, p_e, k)$\;
        $\text{result noisy probabilistic ending}$ $\gets$ Axelrod.tournament$(\text{players}, p_n, p_e, k)$\;

@@

A total of \uniquenumberofseeds trials of Algorithm~\ref{algorithm:data_generation} have been
run. For each trial the results for 4 different tournaments were collected,
thus a total of \numberofalltournaments $(\uniquenumberofseeds \times 4)$ tournament results have been
retrieved. Each tournament outputs a result summary in the form of
Table~\ref{table:output_result}.

The result summary, Table~\ref{table:output_result}, has \(N\) number of rows
because each row contains information for each strategy that participated in the
tournament. The information includes the strategy's rank, median score, the rate
with which the strategy cooperated $(C_r)$, its match win count and the probability that
the strategy cooperated in the opening move. Moreover, the probabilities of a strategy
being in any of the four states ($CC, CD, DC, DD$), and the rate of which the
strategy cooperated after each state.

\newcolumntype{g}{>{\columncolor{Gray}}c}
\begin{table}[h!]
    \begin{center}
    \resizebox{\textwidth}{!}{
    \begin{tabular}{ccccccgcgcgcgcg}
    \toprule 
    & & & & & &   \multicolumn{8}{g}{Rates}  \\
    Rank & Name & Median score & Cooperation rating $(C_r)$ & Win & Initial C &
    CC & CD & DC & DD & CC to C & CD to C & DC to C & DD to C \\
    0 &  EvolvedLookerUp2 2 2 & 2.97 & 0.705 & 28.0 & 1.0 & 0.639 & 0.066 & 0.189 &
    0.106 & 0.836 & 0.481 & 0.568 & 0.8 \\
    1 &  Evolved FSM 16 Noise 05 & 2.875 & 0.697 & 21.0 & 1.0 & 0.676 & 
    0.020 & 0.135 & 0.168 & 0.985 & 0.571 & 0.392 & 0.07 \\
    2 & PSO Gambler 1 1 1 & 2.874 & 0.684 &  23.0 &     1.0 &    0.651 &    0.034 &    0.152 &    0.164
    & 1.000 & 0.283 & 0.000 & 0.136 \\
    3 &  PSO Gambler Mem1 &  2.861 &        0.706 &  23.0 &      1.0 &    0.663
    &    0.042 &    0.145 &    0.150 &  1.000 &  0.510 &  0.000 &  0.122 \\
    4 &          Winner12 &  2.835 &        0.682 &  20.0 &      1.0 &
    0.651 &    0.031 &    0.141 &    0.177 &  1.000 &  0.441 &  0.000 &  0.462 \\
    $\dots$ & $\dots$ & $\dots$ & $\dots$ & $\dots$ & $\dots$ & $\dots$ & $\dots$ &
    $\dots$ & $\dots$ & $\dots$ & $\dots$ & $\dots$ & $\dots$ \\
    \bottomrule
    \end{tabular}}
\end{center}
\caption{Output result of a single tournament.}\label{table:output_result}
\end{table}

A measure that has been manually included
is the \textbf{normalised rank}. The normalised rank, denoted as $r$, is
calculated as a strategy's rank divided by the tournament's size ($N$). The
normalised rank will be used in the next section to evaluate the performance of
strategies.

Table
A summary of the number of entries each strategy
has is given by Table.
Each strategy participated on average in 5180 tournaments of each type. The
strategy that participated in the most tournaments () is and the strategy
with the minimum enties is.

\begin{table}[!htbp]
    \begin{center}
    \resizebox{.45\textwidth}{!}{
        \input{summary_table_of_entries.tex}
    }
\end{center}
% \caption{Top performances for each tournament type based on $\bar{r}$.}
% \label{table:top_performances}
\end{table}

@@

This section evaluates the performance of 194 IPD strategies. The performance of
each strategy is evaluated in four tournament types, which were presented in Section
\ref{section:data_collection}, followed by an evaluation of their performance
over all the \numberofalltournaments simulated tournaments of this work.

Each strategy participated in multiple tournaments of the same type (on average 5180).
For example Tit For Tat has participated in a total of 5114
tournaments of each type. The strategy's normalised rank distribution in these
is given in Figure~\ref{fig:tit_for_tat_r_distribution}. A value of \(r =
0\) corresponds to a strategy winning the tournament where a value of
\(r = 1\) corresponds to the strategy coming last. Because of the strategies'
multiple entries their performance is evaluated based on the
\textbf{median normalised rank} denoted as \(\bar{r}\).

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.55\textwidth]{../images/tit_for_tat_r_distributions.pdf}
    \caption{Tit For Tat's $r$ distribution in tournaments. The best performance
    of the strategy has been in standard tournaments where it achieved a $\bar{r}$
    of 0.339.}
    \label{fig:tit_for_tat_r_distribution}
\end{figure}

@@

Figure~\ref{fig:std_results} gives the distributions of $r$ for the top
ranked strategies. The distributions are skewed towards zero and the highest
median is at 0.075. This indicates that the top ranked strategies are very
likely to perform well in a standard tournament despite the tournament size,
the opponents, the turns etc.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.55\textwidth]{../images/performance_standard.pdf}
    \caption{$r$ distributions of top 15 strategies in standard tournaments.}\label{fig:std_results}
\end{figure}

The top strategies in noisy tournaments are shown in Figure~\ref{fig:noisy_results}. These include deterministic strategies, such
as Tit For 2 Tats~\cite{Axelrod1980b}, Hard Tit For 2 Tats~\cite{Stewart2012}
and Cycler CCCCCD, and strategies which decide their actions based on the
cooperations to defections ratio, such as ShortMem~\cite{Carvalho2013}, Grumpy,
$e$, $\pi$ and $\phi$ (all from~\cite{axelrodproject}). The Retaliate and
Limited Retaliate strategies are implemented in~\cite{axelrodproject} by the
same contributor. The strategies are designed to defect if the opponent has
tricked the strategy more often than \(x\%\) of the times that the strategy has
done the same. Finally, in $4^{\text{th}}$ and $9^{\text{th}}$ are Hunter
strategies which trying to extort, equivalently, strategies that play cyclically
and defectors.

From Figure~\ref{fig:noisy_results}, it is evident that the normalised rank
distributions in noisy environments are more variant with higher medians
compared to standard tournaments. The distributions are bimodal.
This indicates that although the top ranked strategies mainly performed
well, there are several tournaments that they ranked in the bottom half.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.55\textwidth]{../images/performance_noise.pdf}
    \caption{\(r\) distributions for best performed strategies in noisy tournaments.}
    \label{fig:noisy_results}
\end{figure}

The 15 top ranked strategies in probabilistic ending tournaments include
Fortress 3, Fortress 4 (both introduced in~\cite{Ashlock2006}),
Raider~\cite{Ashlock2014} and Solution B1~\cite{Ashlock2014}, which are
strategies based on finite state automata introduced by Daniel and Wendy
Ashlock. These strategies have been evolved using reinforcement learning, however,
there were trained to maximise their payoffs in tournaments with fixed turns
(150 specifically) and not in probabilistic ending ones. In probabilistic ending
tournaments it appears that the top ranks are mostly occupied by defecting
strategies. These include Better and Better, Gradual Killer, Hard Prober (all
from ~\cite{prison}), Bully (Reverse Tit For Tat)~\cite{Nachbar1992} and
Defector. Thus, it's surprisingly that EasyGo and Fool Me Forever are ranked
$14^{\text{th}}$ and $15^{\text{th}}$. These strategies are actually the same;
they will defect until their opponent defect, then they will cooperate until the
end. Both strategies have repeatedly ranked highly as shown in
Figure~\ref{fig:probend_results}, and there are cases for which they were the
winners of the tournament.

\begin{figure}[h!]
    \centering
    \includegraphics[width=.55\textwidth]{../images/performance_probend.pdf}
    \caption{\(r\) distributions for best performed strategies in probabilistic ending tournaments.}
    \label{fig:probend_results}
\end{figure}

The distributions of the normalised rank in probabilistic ending tournaments,
shown in
Figure~\ref{fig:probend_results}, are less variant than those of noisy
tournaments. The medians are lower than 0.1 and the distributions are skewed
towards 0. Though the large difference between the means and the medians
indicates some outliers, the strategies have overall performed well in the
probabilistic ending tournaments that they participated.

In tournaments with both noise and an unspecified number of turns several of
the top ranked strategies are strategies that were highly ranked in noisy
tournaments. This demonstrates that, $e$, $\pi$,
$\phi$ and the Retaliate family are robust in noisy environments regardless of the
number of turns. However, strategies from the top ranks in probabilistic
ending tournaments did not rank highly here.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.55\textwidth]{../images/performance_probend_noise.pdf}
    \caption{\(r\) distributions for best performed strategies in noisy
    probabilistic ending tournaments.}
    \label{fig:noisy_probend_results}
\end{figure}

Up till now, the performances of the 194 strategies have been evaluated for
individual tournament types. The data set considered in this work, described in
Section~\ref{section:data_collection}, contains a total of \numberofalltournaments
tournament results. For this part of the manuscript the strategies are ranked
based on the median normalised rank they achieved over the entire data set.
The top 15 strategies are given in Table~\ref{table:overall_results}
and their normalised rank distributions are given in Figure~\ref{fig:overall_results}.

@@

The aim of this section is to explore the factors that contribute to a
strategy's successful performance. The factors explored are measures regarding a
strategy's behaviour, alogn with measures regarding the tournaments the
strategies competed in. These are given in Table~\ref{table:manual_measures}.

>>>>

\newcommand{\uniquenumberofseeds}{\input{unique_number_of_seeds.tex}}
\newcommand{\numberofalltournaments}{\input{number_of_all_tournaments.tex}}
\newcommand{\numberofstrategies}{\input{unique_number_of_strategies.tex}}

@@

\begin{abstract}
The Iterated Prisoner's Dilemma has been used for decades as a model of
behavioural interactions. From the celebrated performance of Tit for Tat, to the
introduction of the zero-determinant strategies, to the use of sophisticated
structures such as neural networks, the literature has been exploring the
performance of strategies in the game for years. The results of the literature,
however, have been relying on the performance of specific strategies in a finite
number of tournaments, whereas this manuscript evaluates
\numberofstrategies strategies' effectiveness in more than 40000 tournaments.
The top ranked strategies are presented, and moreover, the impact of features on
their success are analysed using machine learning techniques. The analysis
determines that the cooperating ratio of a
strategy in a given tournament compared to the mean and median cooperator is the
most important feature. The conclusions are distinct for different types of
tournaments. A strategy with a theory of mind would aim to be the mean/median
cooperator in standard tournaments, whereas in tournaments with probabilistic
ending it would aim to cooperate 10\% of the times the median cooperator did.
In tournaments with noise and a probability of ending an effective strategy
cooperates 60\% of the times the median/mean cooperator did, and in tourmanents
with only noise or even in a setting that the tournament can be any of the
above a strategy should aim to play 67\% as a median/mean cooperator.
\end{abstract}

@@

Another set of IPD strategies that have received a lot of attention
are the zero-determinant strategies (ZDs)~\cite{Press2012}. By
forcing a linear relationship between the payoffs ZDs can ensure that they will
never receive less than their opponents. The
American Mathematical Society's news section stated that ``the world of game
theory is currently on fire''.
ZDs are indeed a set of mathematically unique strategies
and robust in pairwise interactions, however, their simplicity and extortionate
behaviour have been tested. In~\cite{Harper2017} a tournament containing over
200 strategies, including ZDs, was ran and none of the them
ranked in top spots. Instead, the top ranked strategies were a set of
trained strategies based on lookup tables~\cite{Axelrod1987}, hidden markov
models~\cite{Harper2017} and finite state automata~\cite{Miller1996}.

Though only a select pieces of work have been discussed, the IPD literature is
rich, and new strategies and competitions are being published every year. The
question, however, still remains the same: what is the best way to play the
game? Compared to other works, whereas a few selected strategies are evaluated
on a small number of tournaments, this manuscript evaluates the performance of \numberofstrategies
strategies in \numberofalltournaments tournaments. These tournaments do not
consist by just standard round robin tournaments, but also by tournaments with noise
and tournaments with a probabilistic ending. The later part of the paper, evaluates
the impact of features on the performance of the strategies using modern
machine learning techniques. These features include measures regarding a
strategy's behaviour and measures regarding the tournaments. The data set used
in this work has been made publicly available~\cite{data} and can be used
for further analysis and insights.

The different tournament types as well as the data collection, which is made
possible due an open source package called Axelrod-Python~\cite{axelrodproject},
are covered in Section~\ref{section:data_collection}.
Section~\ref{section:top_performances}, focuses on the best performing
strategies for each type of tournament and overall.
Section~\ref{section:evaluation_of_performance}, explores the traits which
contribute to good performance, and finally the results are summarised in
Section~\ref{section:conclusion}. This manuscripts uses several parameters.
These are introduced in the following sections, however, the full set of
parameters and their definitions are given in Appendix~\ref{app:parameters}.

@@

For the purposes of this manuscript a data set containing results of IPD
tournaments has been generated and is available at~\cite{data}. This was done using the
open source package Axelrod-Python~\cite{axelrodproject}, and more specifically,
version 3.0.0. Axelrod-Python allows for different types of IPD computer
tournaments to be simulated whilst containing a list of over 180 strategies.
Most of these are strategies described in the literature with a few exceptions
being strategies that have been contributed specifically to the package. This
paper make use of \numberofstrategies strategies implemented in version 3.0.0. A
list of the strategies is given in the Appendix~\ref{app:list_of_players}.
Though Axelrod-Python features several tournament types, this work considers
only standard, noisy, probabilistic ending and noisy probabilistic ending
tournaments.

\textbf{Standard tournaments}, are tournaments similar to that of Axelrod's
in~\cite{Axelrod1980a}. There are \(N\) strategies which all play an iterated
game of \(n\) number of turns against each other. Note that self interactions
are not included. Similarly, \textbf{noisy
tournaments} have \(N\) strategies and \(n\) number of turns, but at each turn
there is a probability \(p_n\) that a player's action will be flipped.
\textbf{Probabilistic ending tournaments}, are of size \(N\) and after each turn
a match between strategies ends with a given probability \(p_e\). Finally,
\textbf{noisy probabilistic ending} tournaments have both a noise probability
\(p_n\) and an ending probability \(p_e\). For smoothing the simulated results a
tournament is repeated for \(k\) number of times. This was allowed to vary 
in order to evaluate the effect of smoothing. The winner of each tournament
is based on the average score a strategy achieved and not by the number of wins.

The process of collecting tournament results implemented in this manuscript is described by
Algorithm~\ref{algorithm:data_generation}. For each trial a random size \(N\) is
selected, and from the \numberofstrategies strategies a random list of \(N\) strategies is
chosen. For the given list of strategies a standard, a noisy, a probabilistic
ending and a noisy probabilistic ending tournament are performed and repeated
\(k\) times. The parameters for the tournaments, as well as the number of
repetitions, are selected once for each trial. The parameters and their
respective minimum and maximum values are given by
Table~\ref{table:parameters_values}.

@@

The source code for the data collection, as well as the source code for
the analysis, which will be discussed in the following sections, have been written
following best practices~\cite{Aberdour2007, Benureau2018}. It has been packaged
and is available here. %TODO archive code

\begin{algorithm}[!htbp]
    \setstretch{1.35}
    \ForEach{\text{seed} $\in [0, 11420]$}{
        $N \gets \text{randomly select integer}\in [N_{min}, N_{max}]$\;
        $\text{players} \gets  \text{randomly select $N$ players}$\;
        $k \gets  \text{randomly select integer}\in [k_{min}, k_{max}]$\;
        $n \gets  \text{randomly select integer}\in [n_{min}, n_{max}]$\;
        $p_n \gets  \text{randomly select float}\in [p_{n\, min}, p_{n\, max}]$\;
        $p_e \gets   \text{randomly select float}\in [p_{e\, min}, p_{e\, max}]$\;
        \vspace{0.4cm}
        $\text{result standard}$ $\gets$ Axelrod.tournament$(\text{players}, n, k)$\;
        $\text{result noisy}$ $\gets$ Axelrod.tournament$(\text{players}, n, p_n, k)$\;
        $\text{result probabilistic ending}$ $\gets$ Axelrod.tournament$(\text{players}, p_e, k)$\;
        $\text{result noisy probabilistic ending}$ $\gets$ Axelrod.tournament$(\text{players}, p_n, p_e, k)$\;

@@

A total of \uniquenumberofseeds trials of Algorithm~\ref{algorithm:data_generation} have been
run. For each trial the results for 4 different tournaments were collected,
thus a total of \numberofalltournaments $(\uniquenumberofseeds \times 4)$ tournament results have been
retrieved. Each tournament outputs a result summary in the form of
Table~\ref{table:output_result}. Each strategy have participated on average in
5154 tournaments of each type. The strategy with the maximum participation in each
tournament type is Inverse Punisher with 5639 entries. The strategy with the
minimum entries is EvolvedLookerUp 1 1 1 which was selected in 4693 trials.

The result summary, Table~\ref{table:output_result}, has \(N\) number of rows
because each row contains information for each strategy that participated in the
tournament. The information includes the strategy's rank, median score, the rate
with which the strategy cooperated $(C_r)$, its match win count and the
probability that the strategy cooperated in the opening move. Moreover, the
probabilities of a strategy being in any of the four states ($CC, CD, DC, DD$),
and the rate of which the strategy cooperated after each state. A measure that
has been manually included is the \textbf{normalised rank}. The normalised rank,
denoted as $r$, is calculated as a strategy's rank divided by the tournament's
size ($N$). In the next section the performance of these strategies is evaluated
based on their normalised rank.

\newcolumntype{g}{>{\columncolor{Gray}}c}
\begin{table}[!htbp]
    \begin{center}
    \resizebox{\textwidth}{!}{
    \begin{tabular}{ccccccgcgcgcgcg}
    \toprule 
    & & & & & &   \multicolumn{8}{g}{Rates}  \\
    Rank & Name & Median score & Cooperation rating $(C_r)$ & Win & Initial C &
    CC & CD & DC & DD & CC to C & CD to C & DC to C & DD to C \\
    0 &  EvolvedLookerUp2 2 2 & 2.97 & 0.705 & 28.0 & 1.0 & 0.639 & 0.066 & 0.189 &
    0.106 & 0.836 & 0.481 & 0.568 & 0.8 \\
    1 &  Evolved FSM 16 Noise 05 & 2.875 & 0.697 & 21.0 & 1.0 & 0.676 & 
    0.020 & 0.135 & 0.168 & 0.985 & 0.571 & 0.392 & 0.07 \\
    2 & PSO Gambler 1 1 1 & 2.874 & 0.684 &  23.0 &     1.0 &    0.651 &    0.034 &    0.152 &    0.164
    & 1.000 & 0.283 & 0.000 & 0.136 \\
    3 &  PSO Gambler Mem1 &  2.861 &        0.706 &  23.0 &      1.0 &    0.663
    &    0.042 &    0.145 &    0.150 &  1.000 &  0.510 &  0.000 &  0.122 \\
    4 &          Winner12 &  2.835 &        0.682 &  20.0 &      1.0 &
    0.651 &    0.031 &    0.141 &    0.177 &  1.000 &  0.441 &  0.000 &  0.462 \\
    $\dots$ & $\dots$ & $\dots$ & $\dots$ & $\dots$ & $\dots$ & $\dots$ & $\dots$ &
    $\dots$ & $\dots$ & $\dots$ & $\dots$ & $\dots$ & $\dots$ \\
    \bottomrule
    \end{tabular}}
\end{center}
\caption{Output result of a single tournament.}\label{table:output_result}
\end{table}

@@

This section evaluates the performance of \numberofstrategies IPD strategies. The performance of
each strategy is evaluated in four tournament types, which were presented in Section
\ref{section:data_collection}, followed by an evaluation of their performance
over all the \numberofalltournaments simulated tournaments of this work.

Each strategy participated in multiple tournaments of the same type (on average 5154).
For example Tit For Tat has participated in a total of 5114
tournaments of each type. The strategy's normalised rank distribution in these
is given in Figure~\ref{fig:tit_for_tat_r_distribution}. A value of \(r =
0\) corresponds to a strategy winning the tournament where a value of
\(r = 1\) corresponds to the strategy coming last. Because of the strategies'
multiple entries their performance is evaluated based on the
\textbf{median normalised rank} denoted as \(\bar{r}\).

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.6\textwidth]{../images/tit_for_tat_r_distributions.pdf}
    \caption{Tit For Tat's $r$ distribution in tournaments. The best performance
    of the strategy has been in standard tournaments where it achieved a $\bar{r}$
    of 0.339.}
    \label{fig:tit_for_tat_r_distribution}
\end{figure}

@@

Figure~\ref{fig:std_results} gives the distributions of $r$ for the top
ranked strategies. The distributions are skewed towards zero and the highest
median, of the top 15 strategies, is at 0.075. This indicates that the top ranked
strategies perform well in any given standard tournament, despite the opponents and
the number of turns.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.7\textwidth]{../images/performance_standard.pdf}
    \caption{$r$ distributions of top 15 strategies in standard tournaments.}\label{fig:std_results}
\end{figure}

The top strategies in noisy tournaments are shown in Figure~\ref{fig:noisy_results}. These include deterministic strategies, such
as Tit For 2 Tats~\cite{Axelrod1980b}, Slow Tit For Tats~\cite{axelrodproject}, Hard Tit For 2 Tats~\cite{Stewart2012}
and Cycler CCCCCD, and strategies which decide their actions based on the
cooperations to defections ratio, such as ShortMem~\cite{Carvalho2013}, Grumpy
and $e$~\cite{axelrodproject}. The Retaliate and
Limited Retaliate strategies are implemented in~\cite{axelrodproject} by the
same contributor. They are strategies designed to defect if the opponent has
tricked them more often than \(x\%\) of the times that they have
done the same. Finally, in $4^{\text{th}}$ and $9^{\text{th}}$ place are Hunter
strategies which trying to extort, equivalently, strategies that play cyclically
and defectors.

From Figure~\ref{fig:noisy_results}, it is evident that the normalised rank
distributions in noisy environments are more variant with higher medians
compared to standard tournaments. The distributions are bimodal. This indicates
that although the top ranked strategies mainly performed well, there are several
tournaments that they ranked in the bottom half. To gain a better understanding
of behaviour in noisy tournaments, the \(r\) distribution over the noise
probability \(p_n\) for the top 6 strategies of Figure~\ref{fig:noisy_results},
are given in Figure~\ref{fig:effect_of_noise}.

For \(p_n\) strictly lower than 0.5, Grumpy, Tit For 2 Tats and Slow Tit For Two Tat
have a moderate performing, whereas \(e\), Cycle Hunter and Ricky QLearner are
behaving poorly. At \(p_n=0.5\), all the distributions are bimodal because
with a 0.5 probability of noise any strategies corresponds to a random player.
Interestedly, for a \(p_n\) larger than 0.5 all of the 6 strategies become
successful. Note that a value \(p_n=1\) corresponds to a strategy playing the
exact opposite from what they intended. Thus, the successful strategies in
noisy tournaments presented here, are effective when they are not playing as
themselves. If \(p_n\) was strictly less 0.5 then the top ranked strategies
are different, Table~\ref{table:subset_noise_result}. The top spots are mainly
overtaken by the Meta strategies which are based on teams.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.7\textwidth]{../images/performance_noise.pdf}
    \caption{\(r\) distributions for best performed strategies in noisy tournaments.}
    \label{fig:noisy_results}
\end{figure}

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=\textwidth]{../images/noise_effect.pdf}
    \caption{\(r\) distributions for top 6 strategies in noisy tournaments over
    the probability of noisy ($p_n$).}
    \label{fig:effect_of_noise}
\end{figure}

\begin{table}[!htbp]
    \centering
    \resizebox{.30\textwidth}{!}{
    \input{performance_noise_subset_table.tex}}
    \caption{Top performances in noisy tournaments where \(p_n<0.5\).}\label{table:subset_noise_result}
\end{table}

The 15 top ranked strategies in probabilistic ending tournaments include
Fortress 3, Fortress 4 (both introduced in~\cite{Ashlock2006}),
Raider~\cite{Ashlock2014} and Solution B1~\cite{Ashlock2014}, which are
strategies based on finite state automata introduced by Daniel and Wendy
Ashlock. These strategies have been evolved using reinforcement learning, however,
there were trained to maximise their payoffs in tournaments with fixed turns
(150 specifically) and not in probabilistic ending ones. In probabilistic ending
tournaments it appears that the top ranks are mostly occupied by defecting
strategies. These include Better and Better, Gradual Killer, Hard Prober (all
from ~\cite{prison}), Bully (Reverse Tit For Tat)~\cite{Nachbar1992} and
Defector. Thus, it's surprisingly that EasyGo and Fool Me Forever which are
strategies that will defect until their opponent defect, then they will cooperate
until the end, ranked $14^{\text{th}}$ and $15^{\text{th}}$. Upon inspection,
it was concluded that they are actually the same strategy. This was not known to
the authors at the time of data collection. Figure~\ref{fig:probend_results}
verifies that their perfomance is the same. Both strategies have repeatedly
ranked highly and there are cases for which they were the winners of the
tournament.

\begin{figure}[h!]
    \centering
    \includegraphics[width=.7\textwidth]{../images/performance_probend.pdf}
    \caption{\(r\) distributions for best performed strategies in probabilistic ending tournaments.}
    \label{fig:probend_results}
\end{figure}

The distributions of the normalised rank in probabilistic ending tournaments,
shown in
Figure~\ref{fig:probend_results}, are less variant than those of noisy
tournaments. The medians of the top 15 strategies are lower than 0.1 and the distributions are skewed
towards 0. Though the large difference between the means and the medians
indicates some outliers, the strategies have overall performed well in the
probabilistic ending tournaments that they participated.

The distributions of \(r\) over \(p_e\) for the top 6 strategies are also given
by Figure~\ref{fig:effect_of_probend}. It is shown that the rank of the
strategies is decreasing as the probability of the game ending increases and at
the point of \(p_e=0.1\) the strategies become dominant. All the top ranked
strategies in this setting are defecting strategies, and what is exhibited is
that they do better as the probability of the game ending in the next turn
increases, which is inline with the Folk Theorem~\cite{Fudenberg2009}. Consider
tournaments where the probability of the game ending was less than 0.1, the top
ranked strategies based on \(\bar{r}\) are given
by~Table~\ref{table:subset_probend_result}. The top ranked strategies include
strategies based on teams, trained strategies, Grudger and Spiteful Tit for Tat.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=\textwidth]{../images/folk_theorem.pdf}
    \caption{\(r\) distributions for top 6 strategies in probabilistic ending tournaments
    over $p_e$.}
    \label{fig:effect_of_probend}
\end{figure}

\begin{table}[!htbp]
    \centering
    \resizebox{.30\textwidth}{!}{
    \input{performance_probend_subset_table.tex}}
    \caption{Top performances in probabilistic ending tournaments with \(p_e < 0.1\)}
    \label{table:subset_probend_result}
\end{table}

In tournaments with both noise and an unspecified number of turns several of
the top ranked strategies are strategies that were highly ranked in noisy
tournaments. This demonstrates that, $e$, $\pi$,
$\phi$ ($\pi$, $\phi$ are based on the same approach as $e$) and the Retaliate
family are robust in noisy environments regardless of the
number of turns. However, strategies from the top ranks in probabilistic
ending tournaments did not rank highly here. The distributions of
\(r\) shown in Figure~\ref{fig:noisy_probend_results} have the largest
median values compared to the top rank strategies of the other tournament
types.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.7\textwidth]{../images/performance_probend_noise.pdf}
    \caption{\(r\) distributions for best performed strategies in noisy
    probabilistic ending tournaments.}
    \label{fig:noisy_probend_results}
\end{figure}

The distributions of \(r\) for the tournament types indicate that
for probabilistic ending and standard tournaments successful strategies do exist.
For these settings, the top 15 strategies have frequently ranked in
the top spots with only a few exceptions. Contrarily, it appears that noise
cause variation in the normalised ranks, and the strategies can always guarantee
a spot in the top ranks.

Up till now, the performances of the \numberofstrategies strategies have been evaluated for
individual tournament types. The data set considered in this work, described in
Section~\ref{section:data_collection}, contains a total of \numberofalltournaments
tournament results. For this part of the manuscript the strategies are ranked
based on the median normalised rank they achieved over the entire data set.
The top 15 strategies are given in Table~\ref{table:overall_results}
and their normalised rank distributions are given in Figure~\ref{fig:overall_results}.

@@

The aim of this section is to explore the factors that contribute to a
strategy's successful performance. The factors explored are measures regarding a
strategy's behaviour, along with measures regarding the tournaments the
strategies competed in. These are given in Table~\ref{table:manual_measures}.
